SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-10-24 docs/questioning.tex
ammendments in first two sections

\maketitle
\begin{abstract}
Web application attacks, can have different forms depending on the
execution context of the application and the location
of the programming flaw that leads to the attack.
Some of the most dangerous web application attacks,
like Cross-Site Scripting and {\sc sql} injection,
involve the exploitation of the following weakness:
programmers tend to develop applications that may
accept and process data of uncertain origin without
appropriate validation and / or filtering.
Such attacks have been constantly topping the lists
of the various security bulletin providers despite
the numerous countermeasures that have been proposed
over the past 15 years.

In this paper we first systematize the existing knowledge
regarding the various defense mechanisms by setting up
a model for application attacks that are based on
executable source code injection vulnerabilities. This model
highlights the key weakness behind such attacks
and helps us observe defenses from a common
perspective.

@@

Web application attacks may involve security misconfigurations,
broken authentication and session management, and other issues.
Some of the most dangerous web application attacks though
are associated with the processing of external input data
that may contain malicious code, a vulnerability known as
code injection.

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}
and others. Attacks like these top the vulnerability
lists of numerous bulletin providers like
the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc mitre}\footnote{\url{http://cwe.mitre.org/top25/}}
for several years. Consider the case of {\sc owasp}'s Top
Ten project whose main goal is to raise awareness about
web application security by identifying some of the most
critical risks facing organizations which is referenced by
numerous researchers. In its three consecutive Top Ten lists
(2007, 2010, 2013), different executable source code-driven
injection attacks dominate the top five positions.
This indicates that apart from the fact that malicious
users find new ways to bypass defense
mechanisms by using a variety of techniques despite the numerous
countermeasures that are being introduced.
Note that the number of systems developed to counter {\sc sql}
injection attacks until 2006 was more than twenty~\cite{HVO06}.
Since then the number has doubled.
In our research we focus on the top {\sc tba} systems that counter
application attacks, in terms of citations.
% TODO@dimitro: explain why you leave out {\it SecuriFly}~\cite{MLL05}.

@@

Bratus et al.~\cite{BLSPS11} portray the issue in a generic fashion:
{\it ``unexpected (and unexpectedly powerful) computational models
inside targeted systems, which turn a part of the target into a
so-called ``weird machine" programmable by the attacker via crafted inputs
(a.k.a. ``exploits")"}.
An example of the above definition is the
following: The code fragment below, defines the operation of
addition in the Scheme programming language~\cite{AS96,D09}:

@@

\noindent
Consider the case where instead of a number, a function that leads to an
endless loop is passed as an argument by the user. This will cause
the interpreter to enter
an endless loop and lead to a denial of service.
% Keep in mind that a {\sc cia} does not require a programming
% language with first-class functions.
The intuition here is that {\it ``every application that copies
untrusted input verbatim into an output
program is vulnerable to code injection"}.
Ray and Ligatti~\cite{RL12b} actually proved the above
claim based on formal language theory.

@@

Code injection attacks can be divided into two categories.
The first involves binary code and the second executable source code.
An extensive survey on binary code
injection attacks can be found in reference~\cite{LC03}.
Furthermore, specific advances in exploiting such vulnerabilities
(i.e. ``heap smashing", ``arc injection" and others)
have been presented in reference~\cite{PB04}.
Finally, the countermeasures used to detect
such defects have already been surveyed~\cite{YJP12}
(many of them are also included in a book:~\cite{DKZ12}
-- Section 13.8), thus we do not include them in our work.
In this paper, we focus on the defenses that protect
applications against application attacks based on
the executable source code injection category.

@@

Figure~\ref{fig:taxonomy} presents a taxonomy of executable source
code injection attacks. Such attacks may involve source code, either of a
Domain Specific Language ({\sc dsl})
or a Dynamic Language. Note that binary code injection
attacks can only occur when the target system is implemented in
languages lacking array bounds checking, like
{\sc c} and {\sc c}++. Contrariwise, source code-driven
injection attacks can target applications written in various
programming languages with different features and characteristics.

Application attacks that involve {\sc dsl}s,
constitute an important subset of the code injection problem,
as {\sc dsl}s like {\sc sql} and {\sc xml} play an
significant role in the development of either web or mobile
applications. For example, many applications have interfaces
where a user enters input to interact with the
application's data, thereby interacting with the
underlying {\it Relational Database Management System}
({\sc rdbms}).
This input can become part of an
{\sc sql} query and executed on the target {\sc rdbms}.
A code injection attack that exploits the vulnerabilities of these
interfaces by taking advantage of input validation issues like
incorrectly passed parameters or incorrect type handling,
is called an {\sc sql} injection attack~\cite{CERT02,MS09,HVO06,SW06}.
By using similar techniques malicious users
can perform other exploits based on {\sc dsl}s,
like {\sc xp}ath~\cite{SW06,CDL07,MKS09} and {\sc xml}~\cite{MSM13}.
With this kind of attacks, a malicious user can view sensitive
information, destroy or modify protected data, or even crash the entire
application.

A recent class of {\sc cia}s involve
dynamic languages like Python, Perl, JavaScript, and {\sc php}~\cite{SFVM09,EWKK09,SMS13}.
JavaScript injection attacks in particular, comprise a wide subset of dynamic
language-driven attacks. Such attacks are manifested when a web application
accepts and redisplays data of uncertain origin without appropriate validation
and filtering. Based on this flaw, an attacker can
manage to inject a script in the JavaScript engine of a browser and alter its
execution flow~\cite{ELX07}. JavaScript injection attacks are considered as a
critical issue in web application security mainly because they are
associated with major vulnerabilities such as:
{\sc xss} attacks~\cite{SG07},
{\sc csrf} attacks~\cite{LZRL09}
and Cross-Channel Scripting
({\sc xcs}) attacks~\cite{W10,BBB09}.

@@

In this section we provide a step-by-step exploitation
model to understand the process of carrying out
application attacks based on executable source code injection
vulnerabilities. Figure~\ref{fig:attacks} presents the steps
of different application attacks. For the most part,
the steps are common for each attack. This is because
they are based on similar attack vectors as we described in
the previous section. We base the analysis of the
existing defenses on this model. Figure~\ref{fig:attacks} also
illustrates in which step the defenses detect corresponding
attacks ({\sc todo}).

An attack starts with the malicious user sending an
request either to the server or to a web user.
In the first case the attacker
could upload data containing a specially crafted script
(in a blog comment) or create a request with
additional {\sc sql} code.

>>>>

\maketitle
\begin{abstract}

  Some of the most dangerous web application attacks, like Cross-Site
  Scripting and {\sc sql} injection, exploit the oversight of
  programmers that develop applications which may accept and process
  data of uncertain origin without appropriate validation and / or
  filtering. Such attacks, relying on source code injection
  vulnerabilities, have been constantly topping the lists of the
  various security bulletin providers despite the numerous
  countermeasures that have been proposed over the past 15 years.

  In this paper we systematize the existing knowledge regarding
  various defense mechanisms against source code injection by setting
  up a model that highlights the key weaknesses behind such attacks and
  helps us observe defenses from a common perspective.

@@

Web application attacks may involve security misconfigurations,
broken authentication and session management, or other issues.
Some of the most dangerous web application attacks, though,
are associated with the processing of external input data
that may contain malicious code, a vulnerability known as
code injection.

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations which
is referenced by numerous researchers. In its three consecutive Top
Ten lists (2007, 2010, 2013), different executable source code-driven
injection attacks dominate the top five positions. Malicious users
find new ways to bypass defense mechanisms by using a variety of
techniques despite the numerous countermeasures that are being
introduced. Note that the number of systems developed to counter {\sc
  sql} injection attacks until 2006 was more than twenty~\cite{HVO06}.
Since then the number has doubled. In our research we focus on the top
{\sc tba} systems that counter application attacks, in terms of
citations.
% TODO@dimitro: explain why you leave out {\it SecuriFly}~\cite{MLL05}.

@@

Bratus et al.~\cite{BLSPS11} portray the issue in a generic fashion:
{\it ``unexpected (and unexpectedly powerful) computational models
inside targeted systems, which turn a part of the target into a
so-called `weird machine' programmable by the attacker via crafted inputs
(a.k.a. `exploits')''}.
An example of the above definition is the
following: The code fragment below defines the operation of
addition in the Scheme programming language~\cite{AS96,D09}:

@@

\noindent
Consider the case where instead of a number, a function that leads to an
endless loop is passed as an argument by the user. This will cause
the interpreter to enter
an endless loop and lead to a denial of service.
% Keep in mind that a {\sc cia} does not require a programming
% language with first-class functions.
The intuition here is that {\it ``every application that copies
untrusted input verbatim into an output
program is vulnerable to code injection''}.
Ray and Ligatti~\cite{RL12b} actually proved the above
claim based on formal language theory.

@@

Code injection attacks can be divided into two categories. The first
involves binary code and the second executable source code. An
extensive survey on binary code injection attacks can be found
in~\cite{LC03}. Specific advances in exploiting such vulnerabilities
(i.e. ``heap smashing", ``arc injection" and others) have been
presented in~\cite{PB04}. The countermeasures used to detect such
defects have already been surveyed~\cite{YJP12} (many of them are also
included in a book:~\cite{DKZ12} -- Section 13.8). In the present work
we do not include binary code injection, focusing instead on defenses
that protect applications against application attacks based on the
executable source code injection category.

@@

Figure~\ref{fig:taxonomy} presents a taxonomy of executable source
code injection attacks. Such attacks may involve source code, either of a
Domain Specific Language ({\sc dsl})
or a Dynamic Language.

Application attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s like {\sc sql} and
{\sc xml} play an significant role in the development of either web or
mobile applications. For example, many applications have interfaces
where a user enters input to interact with the application's data,
thereby interacting with the underlying Relational Database Management
System ({\sc rdbms}). This input can become part of an {\sc sql} query
and executed on the target {\sc rdbms}. A code injection attack that
exploits the vulnerabilities of these interfaces by taking advantage
of input validation issues like incorrectly passed parameters or
incorrect type handling, is called an {\sc sql} injection
attack~\cite{CERT02,MS09,HVO06,SW06}. By using similar techniques
malicious users can perform other exploits based on {\sc dsl}s, like
{\sc xp}ath~\cite{SW06,CDL07,MKS09} and {\sc xml}~\cite{MSM13}. With
this kind of attacks, a malicious user can view sensitive information,
destroy or modify protected data, or even crash the entire
application.

A recent class of {\sc cia}s involve dynamic languages like Python,
Perl, JavaScript, and {\sc php}~\cite{SFVM09,EWKK09,SMS13}. JavaScript
injection attacks, in particular, make up a large subset of dynamic
language-driven attacks. Such attacks are manifested when a web
application accepts and redisplays data of uncertain origin without
appropriate validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}. JavaScript injection
attacks are considered as a critical issue in web application security
mainly because they are associated with major vulnerabilities such as:
{\sc xss} attacks~\cite{SG07}, {\sc csrf} attacks~\cite{LZRL09} and
Cross-Channel Scripting ({\sc xcs}) attacks~\cite{W10,BBB09}.

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
executable source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we described in
the previous section. We base the analysis of the existing defenses on
this model. Figure~\ref{fig:attacks} also illustrates in which step
the defenses detect corresponding attacks ({\sc todo}).

An attack starts with the malicious user sending an request either to
the server or to a web user. In the first case the attacker could
upload data containing a specially crafted script (for example, in a
blog comment) or create a request with additional {\sc sql} code.
